### 2019-01-04 notebook/fuzzy-matching.Rmd
notebook

---
title: "fuzzy-matching.Rmd"
author: "Carl Boettiger"
date: "1/3/2019"
output: html_document
---

```{r}
library(tidyverse)
library(taxadb)
```

```{r}
td_create("itis")
```

```{r}
url <- "ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/SpeciesList.txt"
col_names <- c("Seq", "AOU", "English_Common_Name", "French_Common_Name",
               "Spanish_Common_Name", "order", "family", "genus", "species")
bbs <- readr::read_table(url, 
                         skip = 9, 
                         col_names = col_names, 
                         col_types = "ccccccccc")
bbs <- bbs %>% mutate(species = paste(genus, species))
#readr::write_tsv(bbs, "inst/extdata/bbs.tsv")

```

```{r}
name <- bbs$species
authority <- "itis"
db <- td_connect()
match <- "contains"
```

# Strategy 1: SQL-based fuzzy match

```{r}
name_pattern <- switch(match,
                         starts_with = paste0(name, "%"),
                         contains =  paste0("%", name, "%")
  )

  system.time({
  out <- purrr::map_dfr(name_pattern,
          function(pattern)
            taxa_tbl(authority, "taxonid", db) %>%
            filter(name %like% pattern) %>% collect()
         )
  })
out
```

# Strategy 2: Extract a smaller table

```{r}
  ## Strategy: extract all potential matches by Genus alone.  assumes first name is a genus name!
  only_genus <- function(name)  stringi::stri_extract_first_words(name)
  id_tbl <- ids(only_genus(name), authority = authority, db = db, collect = FALSE) %>%
    select(name) %>%
    inner_join(select(taxa_tbl(authority, "hierarchy"), id, genus), by = c(name = "genus")) %>%
    select(id) %>% inner_join(taxa_tbl(authority, "taxonid"), by = "id") %>%
    distinct() %>%
    collect()

  name_regex <- switch(match,
                         starts_with = paste0(name, ".*"),
                         contains =  paste0(".*", name, ".*")
  )

  id_tbl <- collect(taxa_tbl(authority, "taxonid"))

  ## Using the genus subset -- a much smaller list of matches -- is this good or bad?
  system.time({
    out2 <- purrr::map_dfr(name_regex, function(pattern)
      filter(id_tbl, grepl(pattern, name))
    )
  })

```

```{r}   
  ## In memory, even slower!! 
  system.time({
    id_tbl <-  collect(taxa_tbl(authority, "taxonid"))
    out2 <- purrr::map_dfr(name_regex, function(pattern)
      filter(id_tbl, grepl(pattern, name))
    )
  })

```