### 2019-02-12 notebook/algae-explore.Rmd
starting to poke around algae names

```{r}
library(tidyverse)
library(taxadb)
```

Populate the `taxadb` database (only required right after installing `taxadb`, otherwise this will just note the previously installed tables).

```{r}
td_create("all")
```

I'm using names given for Red, Green, and Brown Algae from [Guiry (2012)](https://doi.org/10.1111/j.1529-8817.2012.01222.x):

Open Tree Taxonomy (OTT) has 32,347 names recognized as belonging to one the three phyla:

```{r}
ott_phyla <- bind_rows(
  descendants(name = "Cyanobacteria", rank = "phylum", authority = "ott"),
  descendants(name = "Rhodophyta", rank = "phylum", authority = "ott"),
  descendants(name = "Phaeophyceae", rank = "phylum", authority = "ott")
)

ott_phyla
```

GBIF has 7,412 recognized names

```{r}
gbif_phyla <- bind_rows(
  descendants(name = "Cyanobacteria", rank = "phylum", authority = "gbif"),
  descendants(name = "Rhodophyta", rank = "phylum", authority = "gbif"),
  descendants(name = "Phaeophyceae", rank = "phylum", authority = "gbif")
)

gbif_phyla
```

How many GBIF names also match the names given in OTT? Looks like only 2,923 exact matches.

```{r}
gbif_in_ott <- gbif_phyla %>% 
  select(gbif_id = taxonID, scientificName, taxonRank) %>%
  inner_join(ott_phyla)
```